---
title: "Modelling aperiodic activity"
author: "Zachariah R Cross"
date: "2023-10-27"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Load packages

```{r setup, include=FALSE}

# apply global settings
knitr::opts_chunk$set(echo = TRUE)
options(scipen=999)

# load packages
library(tidyverse)
library(Rmisc)
library(RColorBrewer)
library(hrbrthemes)
library(sjmisc)
library(sjlabelled)
library(lme4)
library(lmerOut)
library(lmerTest)
library(car)
library(ggeffects)
library(performance)
library(ggpubr)
library(rempsyc)
library(ggExtra)
library(broom)
library(plotly)
library(emmeans)
library(purrr)
library(rempsyc)
library(ggseg)
library(ggsegDKT)
library(forcats)
library(ggridges)
library(viridis)
library(splines)

# set contrasts
options(contrasts=c("contr.Sum","contr.Helmert"))
options(decorate.contr.Sum="",decorate.contrasts=c("[","]"))

# functioning for modifying tick labels
number_ticks <- function(n) {function(limits) pretty(limits, n)}

```

## Read in data

```{r load data}

average_df <- read.csv("aperiodic_slope_models.csv") %>% 
  select(-X)

```

# ridge line plots to show distribution

```{r ridgeline}

# Convert DKT to factor with reverse alphabetical order
average_df$DKT <- factor(average_df$DKT, levels = rev(unique(average_df$DKT[order(average_df$DKT)])))

# Plot with reverse alphabetically ordered DKT
ggplot(average_df, aes(x = Slope, y = DKT, fill = ..x..)) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  scale_fill_viridis(name = "Slope", option = "C") +
  labs(y = "", x = "Aperiodic Slope") +
  facet_grid(~condition) +
  theme_minimal() +
  theme(
    panel.spacing = unit(0.1, "lines"),
    strip.text.x = element_text(size = 14, colour = "black"),
    strip.text.y = element_text(size = 14, colour = "black"),
    legend.position = "none",
    legend.spacing.x = unit(0.2, 'cm'),
    legend.key.size = unit(0.4, "cm"),
    legend.background = element_rect(fill = alpha('blue', 0)),
    legend.text = element_text(size = 10, colour = "black"),
    legend.title = element_text(size = 10, colour = "black", face = "bold"),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 14),
    plot.title = element_text(size = 12, face = "bold"))

# generate ridge line plot for offset
ggplot(average_df, aes(x = Offset, y = DKT, fill = ..x..)) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  scale_fill_viridis(name = "Offset", option = "C") +
  labs(y="", x = "Aperiodic Offset") +
  facet_grid(~condition) +
  theme_minimal() +
    theme(legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)) +  
  theme(legend.position="NONE",
        legend.spacing.x = unit(0.2, 'cm'),
        legend.key.size = unit(0.4, "cm"),
        legend.background = element_rect(fill=alpha('blue', 0)),
        legend.text = element_text(size = 10, colour = "black"),
        legend.title = element_text(size = 10, colour = "black", face = "bold"),
        strip.text.x = element_text(size = 14, colour = "black"),
        strip.text.y = element_text(size = 14, colour = "black"),
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        plot.title = element_text(size = 12, face = "bold"))

```

## Get demographic information

```{r data frame aperiodic}

# get number of males/females by condition
sex_condition <- average_df %>%
  ungroup() %>% 
  select(subj, sex, age) %>%
  group_by(sex, subj) %>% 
  dplyr::summarise(age = mean(age)) %>% 
  dplyr::count(sex, sort = TRUE)

# make unique subjects a data frame
subj_list <- unique(average_df$subj)
subj_list <- as.data.frame(subj_list)

# check length of channels
length(unique(average_df$realID))

# get information for age
summary(average_df$age)

# remove subcortical bc not enough observations for model
average_df <- average_df %>% 
  filter(DKT != "subcortical")

```

# Run large mixed model to test hypotheses A and B

```{r lmer analysis}

large_df <- average_df %>%
  filter(DKT != "amygdala" & DKT != "hippocampus") %>%
  select(Slope, Offset, subj, DKT, age, realID) %>%
  mutate(Cortex = case_when(
    DKT %in% c("caudal middle frontal", "insula", "caudal anterior cingulate",
               "fusiform", "inferior temporal", "inferior parietal", 
               "superior temporal", "middle temporal", "parahippocampal", 
               "rostral middle frontal", "posterior cingulate", 
               "superior parietal", "medial orbitofrontal", 
               "inferior frontal", "lateral orbitofrontal", "temporal pole") ~ "Association",
    DKT %in% c("postcentral", "precentral", "lateral occipital") ~ "Sensorimotor",
    TRUE ~ NA_character_  # In case there are other regions not covered
  )) %>%
  group_by(subj, realID, Cortex, DKT) %>%
  dplyr::summarise(Slope = mean(Slope),
                   Offset = mean(Offset),
                   age = mean(age)) %>% 
  na.omit()

large_df$subj <- as.factor(large_df$subj)
large_df$realID <- as.factor(large_df$realID)
large_df$Cortex <- as.factor(large_df$Cortex)
large_df$DKT <- as.factor(large_df$DKT)

# Mixed effects model for slope
large_model_slope <- lmer(Slope ~ ns(age,2) * Cortex + (1|subj) + (1|DKT), data = large_df, control = lmerControl(optimizer = "bobyqa"))
summary(large_model_slope)

# Plot interaction for slope model
data_slope <- ggemmeans(large_model_slope, terms = c("age [all]", "Cortex"),
                     ci.lvl = 0.83) %>% 
  dplyr::rename(Slope = predicted, age = x, Cortex = group)

ggplot(data_slope, aes(age, Slope, color = Cortex)) +
  geom_path(aes()) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = Cortex), alpha = 0.4) +
  #scale_y_continuous(limits = c(1.70, 2.7), breaks = seq(1.70, 2.70, by = .20)) +
  #scale_x_continuous(limits = c(5, 55), breaks = seq(5, 55, by = 5)) +
  ylab("Aperiodic Slope") +
  xlab("Age [years]") +
  scale_color_manual(values = c("turquoise4", "tan1")) + # Adjust colors as needed
  scale_fill_manual(values = c("turquoise4", "tan1")) + # Adjust colors as needed for ribbon
  theme_classic() +  
  theme(legend.position = "top",
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16),
        axis.text.x = element_text(size = 14),
        axis.text.y = element_text(size = 14),
        strip.text.x = element_text(size = 12),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())

# Define the age values you want to compare
age_vals_slope <- seq(5.93, 54, by = 10)

# Run the emmeans function with the specified age values
em_slope <- emmeans(large_model_slope, pairwise ~ age * Cortex, 
                    at = list(age = age_vals_slope), 
                    type = "response", adjust = "none")
em_slope

# Mixed effects model for offset
large_model_offset <- lmer(Offset ~ ns(age,2) * Cortex + (1|subj) + (1|DKT), data = large_df, control = lmerControl(optimizer = "bobyqa"))
summary(large_model_offset)

# extract predicted effects for offset
data_offset <- ggemmeans(large_model_offset, terms = c("age [all]", "Cortex"),
                     ci.lvl = 0.83) %>% 
  dplyr::rename(Offset = predicted, age = x, Cortex = group)

# plot for offset
ggplot(data_offset, aes(age, Offset, color = Cortex)) +
  geom_path(aes()) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = Cortex), alpha = 0.4) +
  #scale_y_continuous(limits = c(1.70, 2.7), breaks = seq(1.70, 2.7, by = .20)) +
  #scale_x_continuous(limits = c(5, 55), breaks = seq(5, 55, by = 5)) +
  ylab("Aperiodic Offset") +
  xlab("Age [years]") +
  scale_color_manual(values = c("turquoise4", "tan1")) + # Adjust colors as needed
  scale_fill_manual(values = c("turquoise4", "tan1")) + # Adjust colors as needed for ribbon
  theme_classic() +  
  theme(legend.position = "top",
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16),
        axis.text.x = element_text(size = 14),
        axis.text.y = element_text(size = 14),
        strip.text.x = element_text(size = 12),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())

```

## Run mixed-effects models and extract predicted effects for slope

```{r lmer loop through each ROI for slope}

# Linear mixed-effects model function for looping over DKT regions
run_lmm_and_extract_effects <- function(data) {
  data <- na.omit(data)  # Filter out rows with missing values
  if (nrow(data) == 0) {
    cat("Skipping: no non-NA data available for this DKT region.\n")
    return(data.frame())  # Return an empty data frame if no data
  }
  
  model <- lmer(Slope ~ age * condition + (1 | subj/realID) + (1|task), data = data,
                control = lmerControl(optimizer = "bobyqa"))
  emm <- ggemmeans(model, terms = c("age [all]", "condition"), ci.lvl = 0.83)
  
  return(as.data.frame(emm))
}

# Create an empty list to populate with DKT labels
DKT_levels <- unique(average_df$DKT)
combined_results <- data.frame()  # Create an empty data frame for combined results
results_df_slope <- data.frame()  # Create an empty data frame for results

# Apply lmer function to each DKT area
for (level in DKT_levels) {
  cat("Processing DKT level:", level, "\n")  # Print the current DKT level
  
  # Subset the data
  subset_data <- average_df[average_df$DKT == level, ]
  
  # Print the number of rows and a sample of the subset data for debugging
  cat("Number of rows in subset_data:", nrow(subset_data), "\n")
  print(head(subset_data))
  
  # Check if there's enough data to run the model
  if (nrow(subset_data) < 5) {  # Threshold can be adjusted as needed
    cat("Skipping:", level, "- Not enough data.\n")
    next  # Skip this iteration if not enough data
  }
  
  # Run lmer model and extract effects
  results <- run_lmm_and_extract_effects(subset_data)
  if (nrow(results) == 0) {
    cat("Skipping:", level, "- No results returned.\n")
    next  # Skip if no results returned
  }
  
  # Add the DKT level as a column in the results
  results$DKT <- level
  # Append results to combined_results
  combined_results <- rbind(combined_results, results)
  
  # Run lmer model for slope results
  model <- lmer(Slope ~ age * condition + (1 | subj/realID) + (1|task), data = subset_data,
                control = lmerControl(optimizer = "bobyqa"))
  
  summary_model <- summary(model)
  
  # Extract estimates, standard errors, and p-values for the fixed effects
  fixed_effects <- summary_model$coefficients[c(2, 3, 4), ]
  
  # Create a new data frame to store the results for slope
  result_row <- data.frame(
    DKT = level,
    Estimate_Age = fixed_effects[1, "Estimate"],
    SE_Age = fixed_effects[1, "Std. Error"],
    PValue_Age = fixed_effects[1, "Pr(>|t|)"],
    Estimate_Condition = fixed_effects[2, "Estimate"],
    SE_Condition = fixed_effects[2, "Std. Error"],
    PValue_Condition = fixed_effects[2, "Pr(>|t|)"],
    Estimate_Interaction = fixed_effects[3, "Estimate"],
    SE_Interaction = fixed_effects[3, "Std. Error"],
    PValue_Interaction = fixed_effects[3, "Pr(>|t|)"]
  )
  
  # Append the results to the main data frame for slope
  results_df_slope <- rbind(results_df_slope, result_row)
}

# Combine the results into a single data frame
combined_results <- combined_results %>% 
  rename(Slope = predicted, age = x, condition = group)

# Generate table of all results for aperiodic slope
slope_table <- results_df_slope %>% 
  dplyr::rename(ROI = DKT, P_Age = PValue_Age, P_Condition = PValue_Condition,
                P_Interaction = PValue_Interaction)

# Apply FDR correction
results_df_slope$p_age_corrected <- p.adjust(results_df_slope$PValue_Age, method = "fdr")
results_df_slope$p_condition_corrected <- p.adjust(results_df_slope$PValue_Condition, method = "fdr")
results_df_slope$p_interaction_corrected <- p.adjust(results_df_slope$PValue_Interaction, method = "fdr")

```

# Plot main effects for slope

```{r slope main effects}

# age data frame for plotting
age_df <- average_df %>% 
  select(subj, DKT, Slope, age) %>% 
  group_by(subj, DKT) %>% 
  summarise(Slope = mean(Slope),
            age = mean(age))

# rostral middle frontal gyrus
ggplot(age_df, aes(x = age, y = Slope)) + 
  geom_point(color = "brown3", alpha = 0.6, size = 3) +  # Change to brick red
  geom_smooth(method = "lm", linetype = "solid", color = "darkgray") +
  geom_ribbon(aes(ymin = ..ymin.., ymax = ..ymax..), 
              stat = "smooth", method = "lm", fill = "darkgray", alpha = 0.1) +
  #scale_y_continuous(limits = c(1, 3), breaks = seq(1, 3, by = .5)) +
  #scale_x_continuous(limits = c(5, 55), breaks = seq(5, 55, by = 10)) +
  scale_color_viridis_c(option = "magma") +
  ylab("Aperiodic Slope") + xlab ("Age [Years]") +
  theme_classic() +
  facet_wrap(~DKT) +
  theme(legend.position = "none",
        legend.spacing.x = unit(0.1, 'cm'),
        legend.key.size = unit(0.5, "cm"),
        legend.background = element_rect(fill = alpha('blue', 0)),
        legend.text = element_text(size = 10, colour = "black"),
        legend.title = element_text(size = 12, colour = "black"),
        strip.text.x = element_text(size = 12, colour = "black"),
        strip.text.y = element_text(size = 12, colour = "black"),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12),
        plot.title = element_text(size = 10, face = "bold"))

# condition data frame for plotting
condition_df <- average_df %>% 
  select(subj, DKT, Slope, age, condition) %>% 
  group_by(subj, DKT, condition) %>% 
  summarise(Slope = mean(Slope),
            age = mean(age))

# raincloud plot for condition main effect
ggplot(condition_df, aes(y = Slope, x = condition, fill = condition)) +
  geom_flat_violin(aes(fill = condition), position = position_nudge(x = .2, y = 0), adjust = 1.5, trim = FALSE, alpha = .5, colour = NA) +
  geom_line(aes(group = condition), alpha = 0.2, colour = "gray48", size = .8, position = position_dodge(0.3)) +
  geom_boxplot(aes(x = condition, y = Slope, fill = condition), outlier.shape = NA, alpha = .5, width = .2, linewidth = .8, position = position_nudge(x = 0, y = 0), colour = "black") +
  scale_y_continuous(limits = c(0.5, 3.5), breaks = seq(0.5, 3.5, by = .5)) +
  facet_wrap(~DKT) +
  ylab("Aperiodic Slope") + xlab("") +
  scale_color_manual(values = c("turquoise4", "tan1")) +
  scale_fill_manual(values = c("turquoise4", "tan1")) +
  theme_bw() +
  theme(legend.position = "NONE",
        legend.spacing.x = unit(0.2, 'cm'),
        legend.key.size = unit(0.4, "cm"),
        legend.background = element_rect(fill = alpha('blue', 0)),
        legend.text = element_text(size = 10, colour = "black"),
        legend.title = element_text(size = 10, colour = "black", face = "bold"),
        strip.text.x = element_text(size = 14, colour = "black"),
        strip.text.y = element_text(size = 14, colour = "black"),
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        plot.title = element_text(size = 12, face = "bold"))

```

# Prepare cortical atlas plots

```{r prepare cortical atlas}

# Load the DKT atlas data
data("dkt")
regions <- dkt[["data"]][["region"]]
regions <- as.data.frame(regions)
regions <- unique(regions)

# Specify the levels of DKT to duplicate and their corresponding new labels
levels_to_duplicate <- c("lateral occipital",         # 1
                         "lateral occipital",         # 2
                         "parahippocampal",           # 3
                         "posterior cingulate",       # 4
                         "precentral",                # 5
                         "inferior frontal",          # 6
                         "inferior frontal",          # 7
                         "inferior frontal",          # 8
                         "lateral occipital",         # 9
                         "superior parietal",         # 10
                         "superior parietal",         # 11
                         "temporal pole",             # 12
                         "superior temporal",         # 13
                         "caudal middle frontal",     # 14
                         "caudal anterior cingulate", # 15
                         "hippocampus",               # 16
                         "amygdala")                  # 17

new_labels <- c("lingual",                            # 1
                "cuneus",                             # 2
                "entorhinal",                         # 3
                "isthmus cingulate",                  # 4
                "paracentral",                        # 5
                "pars opercularis",                   # 6
                "pars triangularis",                  # 7
                "pars orbitalis",                     # 8
                "pericalcarine",                      # 9
                "precuneus",                          # 10
                "supramarginal",                      # 11
                "middle temporal",                    # 12
                "transverse temporal",                # 13
                "superior frontal",                   # 14
                "rostral anterior cingulate",         # 15
                "entorhinal",                         # 16
                "entorhinal")                         # 17

# Create a data frame with levels to duplicate and their corresponding new labels
df_labels <- tibble(DKT = levels_to_duplicate, new_label = new_labels)

# Duplicate the entire data frame
df_duplicated_slope <- bind_rows(replicate(3, results_df_slope, simplify = FALSE))

# Filter rows that need duplication, adjust DKT column and other columns accordingly
df_duplicated_slope <- df_duplicated_slope %>%
  mutate(duplicate = row_number()) %>%
  left_join(df_labels, by = "DKT") %>%
  mutate(DKT = ifelse(!is.na(new_label), new_label, DKT)) %>%
  select(-new_label) %>%
  filter(!is.na(duplicate)) %>%
  group_by(duplicate) %>%
  slice(rep(1:n(), each = 3)) %>%
  ungroup() %>%
  select(-duplicate)

# Combine with the original data frame and remove duplicates
df_final_slope <- bind_rows(df_duplicated_slope, results_df_slope)
df_final_slope <- distinct(df_final_slope)

# convert data frame to tibble and remove labels which are not in DKT atlas
slope_tibble <- tibble(df_final_slope) %>%
  select(DKT, PValue_Interaction, Estimate_Interaction) %>% 
  filter(DKT != "inferior frontal" & DKT != "temporal pole" & DKT != "subcortical") %>%
  rename(region = DKT, p = PValue_Interaction, B = Estimate_Interaction)

# Apply FDR correction
slope_tibble$p_corrected <- p.adjust(slope_tibble$p, method = "fdr")

```

# Generate cortical atlas plots

```{r cortical atlas plots}

# same thing but with continuous scale
ggplot(slope_tibble) +
  geom_brain(
    atlas = dkt, 
    position = position_brain(hemi ~ side),
    aes(fill = B, color = "transparent")) +  # Set color to transparent to remove region boundaries
  scale_fill_viridis_c(option = "magma") +  # Use inferno palette
  scale_color_manual(values = "transparent") +  # Set color to transparent
  theme_void() +
  theme(legend.position = "right",
        plot.title = element_text(size=14, face="bold"))

```

# Plot condition x age interaction for slope
## be sure to replace the roi's when running the code

```{r plot slope interaction}

mod_plot <- lmer(Slope ~ age * condition + (1 | subj/realID) + (1|task), 
                 data = filter(average_df, DKT == "rostral middle frontal"),
                control = lmerControl(optimizer = "bobyqa"))
summary(mod_plot)

raw_data_plot <- average_df %>% 
  ungroup() %>% 
  select(subj, DKT, Slope, age, condition) %>% 
  group_by(subj, DKT, condition) %>% 
  summarise_all(funs(mean), na.rm = TRUE)

# Plot interaction
emm_plot_interaction <- ggemmeans(mod_plot, terms = c("age", "condition"),
                     ci.lvl = 0.83) %>% 
  rename(Slope = predicted, age = x, condition = group)

ggplot(emm_plot_interaction, aes(age, Slope, color = condition)) +
  geom_path(aes()) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = condition), alpha = 0.4) +
  geom_point(data = filter(raw_data_plot, DKT == "rostral middle frontal"),
             aes(y = Slope, color = condition), position = position_jitter(width = .20), 
             size = 3, alpha = 0.3) +
  scale_y_continuous(limits = c(1, 3), breaks = seq(1, 3, by = .5)) +
  scale_x_continuous(limits = c(5, 55), breaks = seq(5, 55, by = 10)) +
  ylab("Aperiodic Slope") +
  xlab("Age [years]") +
  scale_color_manual(values = c("grey49", "#c41e3a")) + # Adjust colors as needed
  scale_fill_manual(values = c("grey49", "#c41e3a")) + # Adjust colors as needed for ribbon
  theme_classic() +  
  theme(legend.position = "top",
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16),
        axis.text.x = element_text(size = 14),
        axis.text.y = element_text(size = 14),
        strip.text.x = element_text(size = 12),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())

```

# Aperiodic slope across the cortex

```{r cortex slope}

# Recoded DKT variable with sum-to-zero contrast coding
average_df$DKT_contrast <- contr.sum(average_df$DKT)

# Run mixed-effects model for slope
roi_model <- lmer(Slope ~ DKT + condition + age + (1 | subj/realID) + (1 | task), 
                  data = average_df, control = lmerControl(optimizer = "bobyqa"))
Anova(roi_model)

# Extract predicted effects
emm_roi <- ggemmeans(roi_model, terms = "DKT",
                     ci.lvl = 0.83) %>% 
  dplyr::rename(Slope = predicted, DKT = x)

# Generate average data frame for plotting raw data
roi_average <- average_df %>% 
  ungroup() %>% 
  select(subj, DKT, Slope) %>% 
  group_by(subj, DKT) %>% 
  summarise_all(funs(mean), na.rm = TRUE)

# re-order the levels of DKT
desired_order <- rev(c("caudal anterior cingulate",
                       "inferior frontal",
                       "lateral orbitofrontal",
                       "medial orbitofrontal",
                       "caudal middle frontal",
                       "rostral middle frontal",
                       "precentral",
                       "postcentral",
                       "insula",
                       "subcortical",
                       "inferior parietal",
                       "superior parietal",
                       "inferior temporal",
                       "middle temporal",
                       "superior temporal",
                       "temporal pole",
                       "fusiform",
                       "parahippocampal",
                       "hippocampus",
                       "amygdala",
                       "posterior cingulate",
                       "lateral occipital"
                      ))

# Convert x to a factor with desired order of levels
emm_roi$DKT <- factor(emm_roi$DKT, levels = desired_order)
roi_average$DKT <- factor(roi_average$DKT, levels = desired_order)

# convert to a data frame so then we can plot on cortex
emm_roi <- as.data.frame(emm_roi)
whole_slope <- emm_roi %>% 
  select(DKT, Slope)

# Duplicate the entire data frame to match the DKT levels
df_whole_cortex <- bind_rows(replicate(3, whole_slope, simplify = FALSE))

# Filter rows that need duplication, adjust DKT column and other columns accordingly
df_whole_cortex <- df_whole_cortex %>%
  mutate(duplicate = row_number()) %>%
  left_join(df_labels, by = "DKT") %>%
  mutate(DKT = ifelse(!is.na(new_label), new_label, DKT)) %>%
  select(-new_label) %>%
  filter(!is.na(duplicate)) %>%
  group_by(duplicate) %>%
  slice(rep(1:n(), each = 3)) %>%
  ungroup() %>%
  select(-duplicate)

# Combine with the original data frame and remove duplicates
df_final_slope <- bind_rows(df_whole_cortex, whole_slope)
df_final_slope <- distinct(df_final_slope)

# convert data frame to tibble and remove labels which are not in DKT atlas
slope_tibble <- tibble(df_final_slope) %>%
  filter(DKT != "inferior frontal" & DKT != "subcortical") %>%
  rename(region = DKT)

# plot on the ggseg DKT atlas
ggplot(slope_tibble) +
  geom_brain(
    atlas = dkt, 
    position = position_brain(hemi ~ side),
    aes(fill = Slope, color = "transparent")) +  # Set color to transparent to remove region boundaries
  scale_fill_viridis_c(option = "magma", limits = c(1.8, 2.3)) +  # Use inferno palette with specified limits
  scale_color_manual(values = "transparent") +  # Set color to transparent
  theme_void() +
  theme(legend.position = "right",
        plot.title = element_text(size=14, face="bold"))

```

# Aperiodic offset across the cortex

```{r cortex offset}

# Run mixed-effects model for slope
roi_model <- lmer(Offset ~ DKT + condition + age + (1 | subj/realID) + (1|task), 
                  data = average_df, control = lmerControl(optimizer = "bobyqa"))
Anova(roi_model)

# Extract predicted effects
emm_roi <- ggemmeans(roi_model, terms = "DKT",
                     ci.lvl = 0.83) %>% 
  dplyr::rename(Offset = predicted, DKT = x)

# Generate average data frame for plotting raw data
roi_average <- average_df %>% 
  ungroup() %>% 
  select(subj, DKT, Offset) %>% 
  group_by(subj, DKT) %>% 
  summarise_all(funs(mean), na.rm = TRUE)

# Convert x to a factor with desired order of levels
emm_roi$DKT <- factor(emm_roi$DKT, levels = desired_order)
roi_average$DKT <- factor(roi_average$DKT, levels = desired_order)

# convert to a data frame to plot on cortex
emm_roi <- as.data.frame(emm_roi)
whole_Offset <- emm_roi %>% 
  select(DKT, Offset)

# Duplicate the entire data frame
df_whole_cortex <- bind_rows(replicate(3, whole_Offset, simplify = FALSE))

# Filter rows that need duplication, adjust DKT column and other columns accordingly
df_whole_cortex <- df_whole_cortex %>%
  mutate(duplicate = row_number()) %>%
  left_join(df_labels, by = "DKT") %>%
  mutate(DKT = ifelse(!is.na(new_label), new_label, DKT)) %>%
  select(-new_label) %>%
  filter(!is.na(duplicate)) %>%
  group_by(duplicate) %>%
  slice(rep(1:n(), each = 3)) %>%
  ungroup() %>%
  select(-duplicate)

# Combine with the original data frame and remove duplicates
df_final_Offset <- bind_rows(df_whole_cortex, whole_Offset)
df_final_Offset <- distinct(df_final_Offset)

# convert data frame to tibble and remove labels which are not in DKT atlas
Offset_tibble <- tibble(df_final_Offset) %>%
  filter(DKT != "inferior frontal" & DKT != "temporal pole" & DKT != "subcortical") %>%
  rename(region = DKT)

# plot on ggseg DKT atlas
ggplot(Offset_tibble) +
  geom_brain(
    atlas = dkt, 
    position = position_brain(hemi ~ side),
    aes(fill = Offset, color = "transparent")) +  # Set color to transparent to remove region boundaries
  scale_fill_viridis_c(option = "magma", limits = c(6.1, 7.8)) +  # Use inferno palette
  scale_color_manual(values = "transparent") +  # Set color to transparent
  theme_void() +
  theme(legend.position = "right",
        plot.title = element_text(size=14, face="bold"))

```
