---
title: "Modelling memory and aperiodic activity"
author: "Zachariah R Cross"
date: "2023-11-01"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Load packages

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

# load packages
library(tidyverse)
library(Rmisc)
library(RColorBrewer)
library(hrbrthemes)
library(sjmisc)
library(sjlabelled)
library(lme4)
library(lmerOut)
library(lmerTest)
library(car)
library(splines)
library(ggeffects)
library(performance)
library(ggpubr)
library(rempsyc)
library(ggExtra)
library(broom)
library(plotly)
library(emmeans)
library(purrr)
library(ggseg)
library(ggsegDKT)

# set contrasts
options(contrasts=c("contr.Sum","contr.Helmert"))
options(decorate.contr.Sum="",decorate.contrasts=c("[","]"))

# functioning for modifying tick labels
number_ticks <- function(n) {function(limits) pretty(limits, n)}

# set jitter width
pd_2 <- position_dodge(width = .5)

```

# Outlier detection and raincloud plot function

```{r outlier detection function}

# outlier detection function
outlierKD <- function(dt, var) {
  var_name <- eval(substitute(var),eval(dt))
  tot <- sum(!is.na(var_name))
  na1 <- sum(is.na(var_name))
  m1 <- mean(var_name, na.rm = T)
  par(mfrow=c(2, 2), oma=c(0,0,3,0))
  boxplot(var_name, main="With outliers")
  hist(var_name, main="With outliers", xlab=NA, ylab=NA)
  outlier <- boxplot.stats(var_name)$out
  mo <- mean(outlier)
  var_name <- ifelse(var_name %in% outlier, NA, var_name)
  boxplot(var_name, main="Without outliers")
  hist(var_name, main="Without outliers", xlab=NA, ylab=NA)
  title("Outlier Check", outer=TRUE)
  na2 <- sum(is.na(var_name))
  message("Outliers identified: ", na2 - na1, " from ", tot, " observations")
  message("Proportion (%) of outliers: ", (na2 - na1) / tot*100)
  message("Mean of the outliers: ", mo)
  m2 <- mean(var_name, na.rm = T)
  message("Mean without removing outliers: ", m1)
  message("Mean if we remove outliers: ", m2)
  response <- readline(prompt="Do you want to remove outliers and to replace with NA? [yes/no]: ")
  if(response == "y" | response == "yes"){
    dt[as.character(substitute(var))] <- invisible(var_name)
    assign(as.character(as.list(match.call())$dt), dt, envir = .GlobalEnv)
    message("Outliers successfully removed", "\n")
    return(invisible(dt))
  } else{
    message("Nothing changed", "\n")
    return(invisible(var_name))
  }
}

# Defining the geom_flat_violin function ----
# Note: the below code modifies the
# existing github page by removing a parenthesis in line 50
"%||%" <- function(a, b) {
  if (!is.null(a)) a else b
}
geom_flat_violin <- function(mapping = NULL, data = NULL, stat = "ydensity",
                             position = "dodge", trim = TRUE, scale = "area",
                             show.legend = NA, inherit.aes = TRUE, ...) {
  layer(
    data = data,
    mapping = mapping,
    stat = stat,
    geom = GeomFlatViolin,
    position = position,
    show.legend = show.legend,
    inherit.aes = inherit.aes,
    params = list(
      trim = trim,
      scale = scale,
      ...
    )
  )
}
#' @rdname ggplot2-ggproto
#' @format NULL
#' @usage NULL
#' @export
GeomFlatViolin <-
  ggproto("GeomFlatViolin", Geom,
          setup_data = function(data, params) {
            data$width <- data$width %||%
              params$width %||% (resolution(data$x, FALSE) * 0.9)
            # ymin, ymax, xmin, and xmax define the bounding rectangle for each group
            data %>%
              group_by(group) %>%
              mutate(
                ymin = min(y),
                ymax = max(y),
                xmin = x,
                xmax = x + width / 2
              )
          },
          draw_group = function(data, panel_scales, coord) {
            # Find the points for the line to go all the way around
            data <- transform(data,
                              xminv = x,
                              xmaxv = x + violinwidth * (xmax - x)
            )
            # Make sure it's sorted properly to draw the outline
            newdata <- rbind(
              plyr::arrange(transform(data, x = xminv), y),
              plyr::arrange(transform(data, x = xmaxv), -y)
            )
            # Close the polygon: set first and last point the same
            # Needed for coord_polar and such
            newdata <- rbind(newdata, newdata[1, ])
            ggplot2:::ggname("geom_flat_violin", GeomPolygon$draw_panel(newdata, panel_scales, coord))
          },
          draw_key = draw_key_polygon,
          default_aes = aes(
            weight = 1, colour = "grey20", fill = "white", size = 0.5,
            alpha = NA, linetype = "solid"
          ),
          required_aes = c("x", "y")
  )

```

## Read in data frames
## read in both the aperiodic and memory data frames and join them

```{r load data}

# load in main data frame
df <- read.csv("aperiodic_final_updated.csv", header = T)

average_df_subj <- df %>% 
  filter(condition == "Rest")
heh <- as.data.frame(unique(average_df_subj$subj))

# load in memory data
memory_data <- read.csv("memory_performance.csv")

# join memory performance to aperiodic data frame
df <- left_join(df, memory_data, by = c("subj","task"))

patient_data <- df %>%  
  select(subj, memory, age, task, condition) %>% 
  filter(condition == "Task") %>% 
  group_by(subj, task, condition) %>% 
  summarise(memory = mean(memory),
            age = mean(age)) %>% 
  ungroup() %>% 
    select(-task, -condition) %>% 
  na.omit()
patient_data$condition <- "pt"

# ensure our factors are indeed factors
patient_data$subj <- as.factor(patient_data$subj)
patient_data$condition <- as.factor(patient_data$condition)

# lets get the mean and standard deviation of performance
descriptive_memory <- patient_data %>% 
  summarise(average = mean(memory),
            sd = sd(memory))

# also grab the range
summary(patient_data$memory)

```

# generate memory performance plot

```{r memory performance}

# load in the control data
control_data <- read.csv("memory_behav_controls.csv")

# ensure our factors are indeed factors
control_data$subj <- as.factor(control_data$subj)
control_data$condition <- as.factor(control_data$condition)

# join the two data frames together
behav_plot_df <- rbind(patient_data, control_data) %>% 
  filter(memory > 0)

# re-label group levels
behav_plot_df$condition <- factor(behav_plot_df$condition,levels=c("ctrl","pt"),
                              labels=c("Control","Patient"))

# let's get the list of the participants in which we have memory performance
subj_list <- as.data.frame(unique(patient_data$subj))

# run linear regression with two cubic splines for patient data only
mem.age <- lm(memory ~ ns(age,2), data = filter(behav_plot_df, condition == "Patient"))
summary(mem.age)

# run mixed model comparing memory against zero
m1. <- lmer(memory ~ 1 + (1 | subj), data = filter(behav_plot_df, condition == "Patient"),
             control = lmerControl(optimizer = "bobyqa"))
summary(m1.)

# plot the results
mem.age.df <- ggemmeans(mem.age,
                     terms = c("age [all]"),
                     ci.lvl = 0.83) %>% 
  dplyr::rename(age = x, memory = predicted, condition = group)

ggplot(mem.age.df, aes(age, memory)) +
  geom_path(aes(colour = condition)) +
  scale_linetype_manual(values=c("solid", "longdash", "solid")) +
  scale_fill_manual(values = c('turquoise4', 'turquoise4', 'turquoise4')) +
  geom_ribbon(aes(fill = condition, ymin = conf.low, ymax = conf.high), alpha = 0.3) +
  scale_x_continuous(limits = c(5, 55), breaks = seq(5, 55, by = 10)) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 6)) +
  labs(y="Recognition Accuracy", x = "Age") +
    geom_hline(yintercept=0, linetype="solid", 
                color = "grey", size=0.5) +
  scale_shape_manual(values=c(1,19,1))+
  scale_color_manual(values=c('#101820FF','grey','turquoise4'))+
  geom_point(aes(shape=condition, color=condition, fill=condition), alpha = 0.6, 
             size = 3, data = behav_plot_df) +
  theme_classic() +  
  theme(legend.position="none",
        axis.title.x = element_text(size=16),
        plot.title = element_text(size=14, face="bold"),
        axis.title.y = element_text(size=16),
        axis.text.x = element_text(size=16),
        axis.text.y = element_text(size=16),
        strip.text.x = element_text(size = 17),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank())

```

# Prepare data frame for analysis
## filter data frames and create resting EEG data frame with memory performance
## also remove outliers

```{r data frame aperiodic}

# average over epochs and filter channels with a model fit of >.90
average_df <- df %>% 
  dplyr::select(subj, age, task, Slope, R.2, DKT,
                Offset, memory) %>%
  filter(task != "Rest") %>% 
  group_by(subj, task, DKT) %>%
  dplyr::summarise(Slope = mean(Slope),
                   Offset = mean(Offset),
            age = mean(age),
            R.2 = mean(R.2),
            memory = mean(memory)) %>% 
  na.omit()

# factorise variables
average_df$subj <- factor(average_df$subj)
average_df$task <- factor(average_df$task)
average_df$DKT <- factor(average_df$DKT)

# create data frame to isolate resting-state
resting_state_df <- df %>% 
  dplyr::select(subj, age, condition, Slope, R.2, DKT, Offset, task) %>%
  filter(task == "Rest" & condition != "Task") %>% 
  group_by(subj, condition, DKT, task) %>%
  dplyr::summarise(Slope = mean(Slope),
                   Offset = mean(Offset),
            age = mean(age),
            R.2 = mean(R.2)) %>% 
  na.omit()

# same as above
memory_df <- average_df %>% 
  select(subj, memory, age) %>% 
  group_by(subj) %>% 
  summarise(memory = mean(memory),
            age = mean(age))

# join memory performance to resting-state slope
memory_rest <- left_join(memory_df, resting_state_df, 
                         by = c("subj", "age")) %>% 
  select(-condition, -R.2) %>% 
  na.omit()
memory_rest$subj <- as.factor(memory_rest$subj)
memory_rest$DKT <- as.factor(memory_rest$DKT)

overall_memory <- average_df %>% 
  select(subj, age, memory, task) %>% 
  group_by(subj, task) %>% 
  dplyr::summarise(memory = mean(memory),
            age = mean(age))

mean_memory <- mean(overall_memory$memory)
sd_memory <- sd(overall_memory$memory)

# run model to examine memory performance
mem_model <- lm(memory ~ age,
                  data = overall_memory)
summary(mem_model)

# remove outliers for slope and offset
# task-related data
outlierKD(average_df, Offset)
outlierKD(average_df, Slope)
# resting-state data
outlierKD(memory_rest, Offset)
outlierKD(memory_rest, Slope)

```

## Run regression models and extract predicted effects for slope
## loop over each DKT region for task-related aperiodic slope and save results

```{r lmer loop through each ROI}

# Linear mixed-effects model function for looping over DKT regions
run_lmm_and_extract_effects <- function(data) {
  data <- na.omit(data)  # Filter out rows with missing values
  model <- lm(memory ~ Slope * age, data = data)
  emm <- ggemmeans(model, terms = c("Slope", "age"),
                     ci.lvl = 0.83)
  
  return(as.data.frame(emm))
}

# Create an empty list to populate with DKT labels
DKT_levels <- unique(average_df$DKT)
combined_results <- data.frame()  # Create an empty data frame for combined results
results_df_slope <- data.frame()  # Create an empty data frame for results

# Apply lmer function to each DKT area
for (level in DKT_levels) {
  cat("Processing DKT level:", level, "\n")  # Print the current DKT level
  
  # Subset the data
  subset_data <- average_df[average_df$DKT == level, ]
  # Run lmer model and extract effects
  results <- run_lmm_and_extract_effects(subset_data)
  # Add the DKT level as a column in the results
  results$DKT <- level
  # Append results to combined_results
  combined_results <- rbind(combined_results, results)
  
  # Run lmer model for slope results
  model <- lm(memory ~ Slope * age, data = subset_data)
  
  summary_model <- summary(model)
  
  # Extract estimates, standard errors, and p-values for the fixed effects
  fixed_effects <- summary_model$coefficients[c(2, 3, 4), ]
  
  # Create a new data frame to store the results for slope
  result_row <- data.frame(
    DKT = level,
    Estimate_Slope = fixed_effects[1, "Estimate"],
    SE_Slope = fixed_effects[1, "Std. Error"],
    PValue_Slope = fixed_effects[1, "Pr(>|t|)"],
    Estimate_Age = fixed_effects[2, "Estimate"],
    SE_Age = fixed_effects[2, "Std. Error"],
    PValue_Age = fixed_effects[2, "Pr(>|t|)"],
    Estimate_Interaction = fixed_effects[3, "Estimate"],
    SE_Interaction = fixed_effects[3, "Std. Error"],
    PValue_Interaction = fixed_effects[3, "Pr(>|t|)"]
  )
  
  # Append the results to the main data frame for slope
  results_df_slope <- rbind(results_df_slope, result_row)
}

# Combine the results into a single data frame
combined_results <- combined_results %>% 
  rename(memory = predicted, Slope = x, age = group)

# generate table of all results for aperiodic slope
slope_table <- results_df_slope %>% 
  dplyr::rename(ROI = DKT, P_Age = PValue_Age, P_Slope = PValue_Slope,
                P_Interaction = PValue_Interaction)

```

# Plot main effect of slope

```{r plot slope}

ggplot(average_df, aes(x = Slope, y = memory)) + 
  geom_point(color = "brown3", alpha = 0.6, size = 3) +  # Change to brick red
  geom_smooth(method = "lm", linetype = "solid", color = "darkgray") +
  geom_ribbon(aes(ymin = ..ymin.., ymax = ..ymax..), 
              stat = "smooth", method = "lm", fill = "darkgray", alpha = 0.1) +
  facet_wrap(~DKT) +
  scale_color_viridis_c(option = "magma") +
  xlab("Aperiodic Slope") + ylab ("Recognition Accuracy") +
  theme_classic() +
  theme(legend.position = "none",
        legend.spacing.x = unit(0.1, 'cm'),
        legend.key.size = unit(0.5, "cm"),
        legend.background = element_rect(fill = alpha('blue', 0)),
        legend.text = element_text(size = 10, colour = "black"),
        legend.title = element_text(size = 12, colour = "black"),
        strip.text.x = element_text(size = 12, colour = "black"),
        strip.text.y = element_text(size = 12, colour = "black"),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12),
        plot.title = element_text(size = 10, face = "bold"))

```

## Run mixed-effects models and extract predicted effects for offset
## loop over each DKT region for task-related aperiodic offset and save results

```{r lmer loop through each ROI}

# Linear mixed-effects model function for looping over DKT regions
run_lmm_and_extract_effects <- function(data) {
  data <- na.omit(data)  # Filter out rows with missing values
  model <- lm(memory ~ Offset * age, data = data)
  emm <- ggemmeans(model, terms = c("Offset", "age"),
                     ci.lvl = 0.83)
  
  return(as.data.frame(emm))
}

# Create an empty list to populate with DKT labels
DKT_levels <- unique(average_df$DKT)
combined_results_offset <- data.frame()  # Create an empty data frame for combined results
results_df_offset <- data.frame()  # Create an empty data frame for results

# Apply lmer function to each DKT area
for (level in DKT_levels) {
  cat("Processing DKT level:", level, "\n")  # Print the current DKT level
  
  # Subset the data
  subset_data <- average_df[average_df$DKT == level, ]
  # Run lmer model and extract effects
  results <- run_lmm_and_extract_effects(subset_data)
  # Add the DKT level as a column in the results
  results$DKT <- level
  # Append results to combined_results
  combined_results_offset <- rbind(combined_results_offset, results)
  
  # Run lmer model for slope results
  model <- lm(memory ~ Offset * age, data = subset_data)
  
  summary_model <- summary(model)
  
  # Extract estimates, standard errors, and p-values for the fixed effects
  fixed_effects <- summary_model$coefficients[c(2, 3, 4), ]
  
  # Create a new data frame to store the results for slope
  result_row <- data.frame(
    DKT = level,
    Estimate_Offset = fixed_effects[1, "Estimate"],
    SE_Offset = fixed_effects[1, "Std. Error"],
    PValue_Offset = fixed_effects[1, "Pr(>|t|)"],
    Estimate_Age = fixed_effects[2, "Estimate"],
    SE_Age = fixed_effects[2, "Std. Error"],
    PValue_Age = fixed_effects[2, "Pr(>|t|)"],
    Estimate_Interaction = fixed_effects[3, "Estimate"],
    SE_Interaction = fixed_effects[3, "Std. Error"],
    PValue_Interaction = fixed_effects[3, "Pr(>|t|)"]
  )
  
  # Append the results to the main data frame for slope
  results_df_offset <- rbind(results_df_offset, result_row)
}

# Combine the results into a single data frame
combined_results_offset <- combined_results_offset %>% 
  rename(memory = predicted, Offset = x, age = group)

# generate table of all results for aperiodic slope
offset_table <- results_df_offset %>% 
  dplyr::rename(ROI = DKT, P_Age = PValue_Age, P_Offset = PValue_Offset,
                P_Interaction = PValue_Interaction)

```

# Plot main effect of offset

```{r plot offset}

ggplot(average_df, aes(x = Offset, y = memory)) + 
  geom_point(color = "dodgerblue3", alpha = 0.6, size = 3) +  # Change to brick red
  geom_smooth(method = "lm", linetype = "solid", color = "darkgray") +
  geom_ribbon(aes(ymin = ..ymin.., ymax = ..ymax..), 
              stat = "smooth", method = "lm", fill = "darkgray", alpha = 0.1) +
  facet_wrap(~DKT) +
  scale_color_viridis_c(option = "magma") +
  xlab("Aperiodic Offset") + ylab ("Recognition Accuracy") +
  theme_classic() +
  theme(legend.position = "none",
        legend.spacing.x = unit(0.1, 'cm'),
        legend.key.size = unit(0.5, "cm"),
        legend.background = element_rect(fill = alpha('blue', 0)),
        legend.text = element_text(size = 10, colour = "black"),
        legend.title = element_text(size = 12, colour = "black"),
        strip.text.x = element_text(size = 12, colour = "black"),
        strip.text.y = element_text(size = 12, colour = "black"),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12),
        plot.title = element_text(size = 10, face = "bold")) 

```

# DKT cortical atlas plots
## relabel our regions to match the ggseg DKT atlas plots
## do this for both the aperiodic slope and offset

```{r prepare cortical atlas}

# Load the DKT atlas data
data("dkt")
regions <- dkt[["data"]][["region"]]
regions <- as.data.frame(regions)
regions <- unique(regions)

# Specify the levels of DKT to duplicate and their corresponding new labels
levels_to_duplicate <- c("lateral occipital",         # 1
                         "lateral occipital",         # 2
                         "parahippocampal",           # 3
                         "posterior cingulate",       # 4
                         "precentral",                # 5
                         "inferior frontal",          # 6
                         "inferior frontal",          # 7
                         "inferior frontal",          # 8
                         "lateral occipital",         # 9
                         "superior parietal",         # 10
                         "superior parietal",         # 11
                         "temporal pole",             # 12
                         "superior temporal",         # 13
                         "caudal middle frontal",     # 14
                         "caudal anterior cingulate", # 15
                         "hippocampus",               # 16
                         "amygdala")                  # 17

new_labels <- c("lingual",                            # 1
                "cuneus",                             # 2
                "entorhinal",                         # 3
                "isthmus cingulate",                  # 4
                "paracentral",                        # 5
                "pars opercularis",                   # 6
                "pars triangularis",                  # 7
                "pars orbitalis",                     # 8
                "pericalcarine",                      # 9
                "precuneus",                          # 10
                "supramarginal",                      # 11
                "middle temporal",                    # 12
                "transverse temporal",                # 13
                "superior frontal",                   # 14
                "rostral anterior cingulate",         # 15
                "entorhinal",                         # 16
                "entorhinal")                         # 17

# Create a data frame with levels to duplicate and their corresponding new labels
df_labels <- tibble(DKT = levels_to_duplicate, new_label = new_labels)

# Duplicate the entire data frame
df_duplicated_slope <- bind_rows(replicate(3, results_df_slope, simplify = FALSE))

# Filter rows that need duplication, adjust DKT column and other columns accordingly
df_duplicated_slope <- df_duplicated_slope %>%
  mutate(duplicate = row_number()) %>%
  left_join(df_labels, by = "DKT") %>%
  mutate(DKT = ifelse(!is.na(new_label), new_label, DKT)) %>%
  select(-new_label) %>%
  filter(!is.na(duplicate)) %>%
  group_by(duplicate) %>%
  slice(rep(1:n(), each = 3)) %>%
  ungroup() %>%
  select(-duplicate)

# Combine with the original data frame and remove duplicates
df_final_slope <- bind_rows(df_duplicated_slope, results_df_slope)
df_final_slope <- distinct(df_final_slope)

# convert data frame to tibble and remove labels which are not in DKT atlas
slope_tibble <- tibble(df_final_slope) %>%
  select(DKT, PValue_Interaction, Estimate_Interaction) %>% 
  filter(DKT != "inferior frontal" & DKT != "temporal pole" & DKT != "subcortical") %>%
  rename(region = DKT, p = PValue_Interaction, B = Estimate_Interaction)

# Apply FDR correction
slope_tibble$p_corrected <- p.adjust(slope_tibble$p, method = "fdr")

# Do the same thing above but for offset
df_duplicated_offset <- bind_rows(replicate(3, results_df_offset, simplify = FALSE))

# Filter rows that need duplication, adjust DKT column and other columns accordingly
df_duplicated_offset <- df_duplicated_offset %>%
  mutate(duplicate = row_number()) %>%
  left_join(df_labels, by = "DKT") %>%
  mutate(DKT = ifelse(!is.na(new_label), new_label, DKT)) %>%
  select(-new_label) %>%
  filter(!is.na(duplicate)) %>%
  group_by(duplicate) %>%
  slice(rep(1:n(), each = 3)) %>%
  ungroup() %>%
  select(-duplicate)

# Combine with the original data frame and remove duplicates
df_final_offset <- bind_rows(df_duplicated_offset, results_df_offset)
df_final_offset <- distinct(df_final_offset)

# convert data frame to tibble and remove labels which are not in DKT atlas
offset_tibble <- tibble(df_final_offset) %>%
  select(DKT, PValue_Offset, Estimate_Offset) %>% 
  filter(DKT != "inferior frontal" & DKT != "subcortical") %>%
  rename(region = DKT, p = PValue_Offset, B = Estimate_Offset)

```

# Generate cortical atlas plots
## here we are plotting the significant interaction (slope x age) on memory for slope and not offset, given that offset did not show any significant interactions

```{r cortical atlas plots}

# generate cortical plot for slope
ggplot(slope_tibble) +
  geom_brain(
    atlas = dkt, 
    position = position_brain(hemi ~ side),
    aes(fill = B, color = "transparent")) +  # Set color to transparent to remove region boundaries
  scale_fill_viridis_c(option = "magma") +  # Adjust scale limits
  scale_color_manual(values = "transparent") +  # Set color to transparent
  theme_void() +
  theme(legend.position = "right",
        plot.title = element_text(size=14, face="bold"))


```

# Plot the interaction for slope
## plot the significant two-way interactions on a scatterplot for slope

```{r plot significant regions slope x age}

ggplot(filter(average_df, DKT == "fusiform"), aes(x=Slope, y=memory, color = age)) + 
  geom_point(alpha = .5, size = 3) +
  geom_smooth(method = "lm", linetype="solid", colour = "gray50") +
  scale_color_gradient(low = "#0091ff", high = "#f0650e", name = "Age [years]") +
  labs(y="Recognition Accuracy", x = "Aperiodic Slope") +
  scale_x_continuous(breaks=number_ticks(6)) +
  scale_y_continuous(breaks=number_ticks(10)) +
  theme_classic()  +
  theme(legend.position="right",
        legend.spacing.x = unit(0.1, 'cm'),
        legend.key.size = unit(0.5, "cm"),
        legend.background = element_rect(fill=alpha('blue', 0)),
        legend.text = element_text(size = 10, colour = "black"),
        legend.title = element_text(size = 12, colour = "black"),
        strip.text.x = element_text(size = 12, colour = "black"),
        strip.text.y = element_text(size = 12, colour = "black"),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12),
        plot.title = element_text(size = 10, face = "bold"))

ggplot(filter(average_df, DKT == "rostral middle frontal"), aes(x=age, y=Slope, color = memory)) + 
  geom_point(alpha = .5, size = 3) +
  geom_smooth(method = "lm", linetype="solid", colour = "gray50") +
  scale_color_gradient(low = "#0091ff", high = "#f0650e", name = "memory") +
  labs(x="Age [years]", y = "Aperiodic Slope") +
  scale_x_continuous(breaks=number_ticks(6)) +
  scale_y_continuous(breaks=number_ticks(10)) +
  theme_classic()  +
  theme(legend.position="right",
        legend.spacing.x = unit(0.1, 'cm'),
        legend.key.size = unit(0.5, "cm"),
        legend.background = element_rect(fill=alpha('blue', 0)),
        legend.text = element_text(size = 10, colour = "black"),
        legend.title = element_text(size = 12, colour = "black"),
        strip.text.x = element_text(size = 12, colour = "black"),
        strip.text.y = element_text(size = 12, colour = "black"),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12),
        plot.title = element_text(size = 10, face = "bold"))

```

# Plot main effect of age (offset)

```{r plot significant offset main effect}

# Plot main effect (rostral middle frontal)
ggplot(filter(average_df, DKT == "rostral middle frontal"), aes(x = Offset, y = memory)) + 
  geom_point(alpha = 0.5, size = 3, color = "#38618c") +
  geom_smooth(method = "lm", linetype = "solid", colour = "#38618c") +
  geom_ribbon(aes(ymin = ..ymin.., ymax = ..ymax..), 
              stat = "smooth", method = "lm", fill = "#38618c", alpha = 0.1) +  # Shade SE in blue
  labs(x = "Aperiodic Offset", y = "Recognition Accuracy") +
  scale_x_continuous(breaks = number_ticks(6)) +
  scale_y_continuous(breaks = number_ticks(10)) +
  theme_classic()  +
  theme(legend.position="right",
        legend.spacing.x = unit(0.1, 'cm'),
        legend.key.size = unit(0.5, "cm"),
        legend.background = element_rect(fill=alpha('blue', 0)),
        legend.text = element_text(size = 10, colour = "black"),
        legend.title = element_text(size = 12, colour = "black"),
        strip.text.x = element_text(size = 12, colour = "black"),
        strip.text.y = element_text(size = 12, colour = "black"),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12),
        plot.title = element_text(size = 10, face = "bold"))

# Plot main effect (inferior parietal)
ggplot(filter(average_df, DKT == "inferior parietal"), aes(x = Offset, y = memory)) + 
  geom_point(alpha = 0.5, size = 3, color = "#38618c") +
  geom_smooth(method = "lm", linetype = "solid", colour = "#38618c") +
  geom_ribbon(aes(ymin = ..ymin.., ymax = ..ymax..), 
              stat = "smooth", method = "lm", fill = "#38618c", alpha = 0.1) +  # Shade SE in blue
  labs(x = "Aperiodic Offset", y = "Recognition Accuracy") +
  scale_x_continuous(breaks = number_ticks(6)) +
  scale_y_continuous(breaks = number_ticks(10)) +
  theme_classic()  +
  theme(legend.position="right",
        legend.spacing.x = unit(0.1, 'cm'),
        legend.key.size = unit(0.5, "cm"),
        legend.background = element_rect(fill=alpha('blue', 0)),
        legend.text = element_text(size = 10, colour = "black"),
        legend.title = element_text(size = 12, colour = "black"),
        strip.text.x = element_text(size = 12, colour = "black"),
        strip.text.y = element_text(size = 12, colour = "black"),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12),
        plot.title = element_text(size = 10, face = "bold"))

# generate cortical plot for offset
ggplot(offset_tibble) +
  geom_brain(
    atlas = dkt, 
    position = position_brain(hemi ~ side),
    aes(fill = B, color = "transparent")) +  # Set color to transparent to remove region boundaries
  #scale_fill_viridis_c(option = "magma") +  # Use inferno palette
  scale_fill_viridis_c(option = "magma") +  # Adjust scale limits
  scale_color_manual(values = "transparent") +  # Set color to transparent
  theme_void() +
  theme(legend.position = "right",
        plot.title = element_text(size=14, face="bold"))

```

# Resting-state memory analysis (slope)

```{r resting state memory}

# Linear mixed-effects model function for looping over DKT regions
run_lmm_and_extract_effects <- function(data) {
  data <- na.omit(data)  # Filter out rows with missing values
  model <- lm(memory ~ Slope * age, data = data)
  emm <- ggemmeans(model, terms = c("Slope", "age"),
                     ci.lvl = 0.83)
  
  return(as.data.frame(emm))
}

# Create an empty list to populate with DKT labels
DKT_levels <- unique(memory_rest$DKT)
combined_results_slope_rest <- data.frame()  # Create an empty data frame for combined results
results_df_slope_rest <- data.frame()  # Create an empty data frame for results

# Apply lmer function to each DKT area
for (level in DKT_levels) {
  cat("Processing DKT level:", level, "\n")  # Print the current DKT level
  
  # Subset the data
  subset_data <- memory_rest[memory_rest$DKT == level, ]
  # Run lmer model and extract effects
  results <- run_lmm_and_extract_effects(subset_data)
  # Add the DKT level as a column in the results
  results$DKT <- level
  # Append results to combined_results
  combined_results_slope_rest <- rbind(combined_results_slope_rest, results)
  
  # Run lmer model for slope results
  model <- lm(memory ~ Slope * age, data = subset_data)
  
  summary_model <- summary(model)
  
  # Extract estimates, standard errors, and p-values for the fixed effects
  fixed_effects <- summary_model$coefficients[c(2, 3, 4), ]
  
  # Create a new data frame to store the results for slope
  result_row <- data.frame(
    DKT = level,
    Estimate_Slope = fixed_effects[1, "Estimate"],
    SE_Slope = fixed_effects[1, "Std. Error"],
    PValue_Slope = fixed_effects[1, "Pr(>|t|)"],
    Estimate_Age = fixed_effects[2, "Estimate"],
    SE_Age = fixed_effects[2, "Std. Error"],
    PValue_Age = fixed_effects[2, "Pr(>|t|)"],
    Estimate_Interaction = fixed_effects[3, "Estimate"],
    SE_Interaction = fixed_effects[3, "Std. Error"],
    PValue_Interaction = fixed_effects[3, "Pr(>|t|)"]
  )
  
  # Append the results to the main data frame for slope
  results_df_slope_rest <- rbind(results_df_slope_rest, result_row)
}

# Combine the results into a single data frame
combined_results_slope_rest <- combined_results_slope_rest %>% 
  rename(memory = predicted, Slope = x, age = group)

# generate table of all results for aperiodic slope
slope_rest_table <- results_df_slope_rest %>% 
  dplyr::rename(ROI = DKT, P_Age = PValue_Age, P_Slope = PValue_Slope,
                P_Interaction = PValue_Interaction)

```

# Plot main effect of resting slope

```{r plot slope resting}

ggplot(memory_rest, aes(x = Slope, y = memory)) + 
  geom_point(color = "brown3", alpha = 0.6, size = 3) +  # Change to brick red
  geom_smooth(method = "lm", linetype = "solid", color = "darkgray") +
  geom_ribbon(aes(ymin = ..ymin.., ymax = ..ymax..), 
              stat = "smooth", method = "lm", fill = "darkgray", alpha = 0.1) +
  #scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = .2)) +
  #scale_x_continuous(limits = c(1.25, 3), breaks = seq(1.25, 3, by = .35)) +
  facet_wrap(~DKT) +
  scale_color_viridis_c(option = "magma") +
  xlab("Aperiodic Slope") + ylab ("Recognition Accuracy") +
  theme_classic() +
  theme(legend.position = "none",
        legend.spacing.x = unit(0.1, 'cm'),
        legend.key.size = unit(0.5, "cm"),
        legend.background = element_rect(fill = alpha('blue', 0)),
        legend.text = element_text(size = 10, colour = "black"),
        legend.title = element_text(size = 12, colour = "black"),
        strip.text.x = element_text(size = 12, colour = "black"),
        strip.text.y = element_text(size = 12, colour = "black"),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12),
        plot.title = element_text(size = 10, face = "bold")) 

```

# Resting-state memory analysis (offset)

```{r resting state memory}
# Linear mixed-effects model function for looping over DKT regions

run_lmm_and_extract_effects <- function(data) {
  data <- na.omit(data)  # Filter out rows with missing values
  model <- lm(memory ~ Offset * age, data = data)
  emm <- ggemmeans(model, terms = c("Offset", "age"),
                     ci.lvl = 0.83)
  
  return(as.data.frame(emm))
}

# Create an empty list to populate with DKT labels
DKT_levels <- unique(memory_rest$DKT)
combined_results_Offset_rest <- data.frame()  # Create an empty data frame for combined results
results_df_Offset_rest <- data.frame()  # Create an empty data frame for results

# Apply lmer function to each DKT area
for (level in DKT_levels) {
  cat("Processing DKT level:", level, "\n")  # Print the current DKT level
  
  # Subset the data
  subset_data <- memory_rest[memory_rest$DKT == level, ]
  # Run lmer model and extract effects
  results <- run_lmm_and_extract_effects(subset_data)
  # Add the DKT level as a column in the results
  results$DKT <- level
  # Append results to combined_results
  combined_results_Offset_rest <- rbind(combined_results_Offset_rest, results)
  
  # Run lmer model for Offset results
  model <- lm(memory ~ Offset * age, data = subset_data)
  
  summary_model <- summary(model)
  
  # Extract estimates, standard errors, and p-values for the fixed effects
  fixed_effects <- summary_model$coefficients[c(2, 3, 4), ]
  
  # Create a new data frame to store the results for Offset
  result_row <- data.frame(
    DKT = level,
    Estimate_Offset = fixed_effects[1, "Estimate"],
    SE_Offset = fixed_effects[1, "Std. Error"],
    PValue_Offset = fixed_effects[1, "Pr(>|t|)"],
    Estimate_Age = fixed_effects[2, "Estimate"],
    SE_Age = fixed_effects[2, "Std. Error"],
    PValue_Age = fixed_effects[2, "Pr(>|t|)"],
    Estimate_Interaction = fixed_effects[3, "Estimate"],
    SE_Interaction = fixed_effects[3, "Std. Error"],
    PValue_Interaction = fixed_effects[3, "Pr(>|t|)"]
  )
  
  # Append the results to the main data frame for Offset
  results_df_Offset_rest <- rbind(results_df_Offset_rest, result_row)
}

# Combine the results into a single data frame
combined_results_Offset_rest <- combined_results_Offset_rest %>% 
  rename(memory = predicted, Offset = x, age = group)

# generate table of all results for aperiodic Offset
Offset_rest_table <- results_df_Offset_rest %>% 
  dplyr::rename(ROI = DKT, P_Age = PValue_Age, P_Offset = PValue_Offset,
                P_Interaction = PValue_Interaction)

```

# Plot main effect of resting offset

```{r plot offset resting}

ggplot(memory_rest, aes(x = Offset, y = memory)) + 
  geom_jitter(width = 6, color = "dodgerblue3", alpha = 0.6, size = 3) +  
  geom_smooth(method = "lm", linetype = "solid", color = "darkgray") +
  geom_ribbon(aes(ymin = ..ymin.., ymax = ..ymax..), 
              stat = "smooth", method = "lm", fill = "darkgray", alpha = 0.1) +
  #scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = .2)) +
  #scale_x_continuous(limits = c(1.25, 3), breaks = seq(1.25, 3, by = .35)) +
  facet_wrap(~DKT) +
  scale_color_viridis_c(option = "magma") +
  xlab("Aperiodic Offset") + ylab("Recognition Accuracy") +
  theme_classic() +
  theme(legend.position = "none",
        legend.spacing.x = unit(0.1, 'cm'),
        legend.key.size = unit(0.5, "cm"),
        legend.background = element_rect(fill = alpha('blue', 0)),
        legend.text = element_text(size = 10, colour = "black"),
        legend.title = element_text(size = 12, colour = "black"),
        strip.text.x = element_text(size = 12, colour = "black"),
        strip.text.y = element_text(size = 12, colour = "black"),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 12),
        plot.title = element_text(size = 10, face = "bold"))

```
